### 2019-01-25 Documents/Latex/Paper.tex
Worked on Praesentation

Obviously, different use cases require vastly different approaches. The most important distinction one needs to make is probably the one between robust solutions that need to produce decent results every run (most end user gaming applications fall under this category), and solutions that are meant to be heavily edited and parameterized and can have some ``bad'' (whatever that means in the respective context) outcomes. Usually, diverse possible results are a key incentive to use procedural generation in the first place, and necessarily the degrees of freedom for the generation system become limited with each additional constraint--especially if those are fixed in code, and not subject to manual assessment and editing.

@@

\subsection{General} \label{cpt:ImplGeneral}
The design idea at the very fundamentals was to implement growth logic on the basis of cells, and understand organisms as a collection of connected cells that interact with each other, but are otherwise mostly independent. Cells then can transform or divide into other cell types, a process which forms complex organisms in the end. This is, in a kind, a derivation from Lindenmayer-Systems and it would indeed be possible to encode all cell states in such a way that it would fulfill the respective definition. Yet the resulting strings would be very long and hard to read, and interactions based on positioning in a 3D space are far from the way L-Systems usually work, so it makes a lot of practical sense to drop the formal constraints of thinking in those limiting terms.

@@

\subsection{Modelled Behaviors} \label{cptr:Influencing}
The following behaviors are simulated:
\begin{itemize}
	\item Plants in water are allowed to have more cells than those without access to it.
	\item Wind/storm blows over a part of the stage, destroying parts of plants that get hit by to much wind.
	\item There is a cone-shaped light source, leading plants to grow in the direction of the light as they divide faster on the light-faced site, have smaller leaves and to have a higher total cell count.
	\item The diameter of cells grows when a higher ``weight'' rest on them, i.e., when they have many attachment descendants.
	\item The growth direction can have a positive or negative correlation with gravity, leading to trees that always grow upward, and trees that ``bend over'', even off the stage.
	\item Cells divisions cannot result in children that collide with external objects or other cells of the plant. Figure~\ref{fig:SelfCollision} shows an example.
		\begin{figure} 
 		 \centering
 		    \includegraphics[width=0.6\textwidth]{SS_SelfCollision.png}
 		 \caption{Self collision enabled (left) and disabled (right).}
 		 \label{fig:SelfCollision}
	\end{figure}
	\item Certain cells---in this case the roots---can grow attached to the ground and up to walls when they hit them (Figure~\ref{fig:RootsOnFloor}).
	\begin{figure} 
 		 \centering
 		    \includegraphics[width=0.6\textwidth]{SS_Roots.png}
 		 \caption{Roots growing on the floor and up a wall.}
 		 \label{fig:RootsOnFloor}
	\end{figure}
\end{itemize}
As hinted above, the numerous collision features of Unreal were used extensively. On the one hand directly to check for overlapping collision shapes (e.g., for the water mechanics) or in the form of ``Raycasts/Line Traces'' that determine where a point that travels on a given line segment hits something. This was, for example, used for the light and wind mechanics, which are modeled by around 10,000 Raycasts each growth iteration.
One possible influence of light is shown in Figure~\ref{fig:GrowToLight}. The light comes from the left, and the source is very close to the tree on the left. This tree has noticeably more leafs and children than the one further in the background, even though the genetics/properties of the plants are the same. Additionally, it can be clearly seen how the tree divided more on the site facing the light source. The image also shows nicely the vastly different possible diameters of cells depending on the number of attachment descendants they have, and the effect of the trees bending upwards as they have a positive correlation with gravity, too.

>>>>

Obviously, different use cases require vastly different approaches. The most important distinction one needs to make is probably the one between robust solutions that need to produce decent results every run (most end user gaming applications fall under this category), and solutions that are meant to be heavily edited and parameterized and can have some ``bad'' (whatever that means in the respective context) outcomes. Usually, diverse possible results are a key incentive to use procedural generation in the first place, and necessarily the degrees of freedom for the generation system become limited with each additional constraint---especially if those are fixed in code, and not subject to manual assessment and editing.

@@

\section{Approach}\label{cpt:ImplGeneral}
The design idea at the very fundamentals was to implement growth logic on the basis of cells, and understand organisms as a collection of connected cells that interact with each other, but are otherwise mostly independent. Cells then can transform or divide into other cell types, a process which forms complex organisms in the end. This is, in a kind, a derivation from Lindenmayer-Systems and it would indeed be possible to encode all cell states in such a way that it would fulfill the respective definition. Yet the resulting strings would be very long and hard to read, and interactions based on positioning in a 3D space are far from the way L-Systems usually work, so it makes a lot of practical sense to drop the formal constraints of thinking in those limiting terms.

@@

\subsection{Modelled Behaviors} \label{cptr:Influencing}
The following behaviors are simulated:
\begin{itemize}
	\item Plants in water are allowed to have more cells than those without access to it.
	\item Wind/storm blows over a part of the stage, destroying parts of plants that get hit by to much wind.
	\item There is a cone-shaped light source, leading plants to grow in the direction of the light as they divide faster on the light-faced site, have smaller leafs and to have a higher total cell count.
	\item The diameter of cells grows when a higher ``weight'' rest on them, i.e., when they have many attachment descendants.
	\item The growth direction can have a positive or negative correlation with gravity, leading to trees that always grow upward, and trees that ``bend over'', even off the stage.
	\item Cells divisions cannot result in children that collide with external objects or other cells of the plant. Figure~\ref{fig:SelfCollision} shows an example.
		\begin{figure} 
 		 \centering
 		    \includegraphics[width=0.6\textwidth]{SS_SelfCollision.png}
 		 \caption{Self collision enabled (left) and disabled (right).}
 		 \label{fig:SelfCollision}
	\end{figure}
	\item Certain cells---in this case the roots---can grow attached to the ground and up to walls when they hit them (Figure~\ref{fig:RootsOnFloor}).
	\begin{figure} 
 		 \centering
 		    \includegraphics[width=0.6\textwidth]{SS_Roots.png}
 		 \caption{Roots growing on the floor and up a wall.}
 		 \label{fig:RootsOnFloor}
	\end{figure}
\end{itemize}
As hinted above, the numerous collision features of Unreal were used extensively. On the one hand directly to check for overlapping collision shapes (e.g., for the water mechanics) or in the form of ``Raycasts/Line Traces'' that determine where a point that travels on a given line segment hits something. This was, for example, used for the light and wind mechanics, which are modeled by around 10,000 Raycasts each growth iteration.
One possible influence of light is shown in Figure~\ref{fig:GrowToLight}. The light comes from the left, and the source is very close to the tree on the left. This tree has noticeably more leafs and children than the one further in the background, even though the genetics/properties of the plants are the same. Additionally, it can be clearly seen how the tree divided more on the site facing the light source. The image also shows nicely the vastly different possible diameters of cells depending on the number of attachment descendants they have, and the effect of the trees bending upwards as they have a positive correlation with gravity, too.
